     ---
title: "Untitled"
author: "Rodney Joujoute"
date: "March 21, 2015"
output: html_document
---

# Course Project Part 2 


This report is for second part of the course project of the Coursera course “Statistical Inference” which is a part of specialization “Data Science”. In this second part, we perform basic inferential analyses using the ToothGrowth data in the R datasets package.

As per the help file of this dataset, the response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

1. Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r dataset, echo=TRUE}
library(datasets)
data(ToothGrowth)

str(ToothGrowth)
```


```{r  rows, echo=TRUE}
nrow(ToothGrowth)
```

```{r factor, echo=TRUE}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

str(ToothGrowth)
```

2. Provide a basic summary of the data.

```{r summary, echo=TRUE}
summary(ToothGrowth)
```

```{r dose, echo=TRUE}
table(ToothGrowth$dose, ToothGrowth$supp)
```

```{r doseplot}
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose))
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.

```{r groupdiff, echo=TRUE }
t.test(len ~ supp, data = ToothGrowth)
```

The p-value is 0.06, and the confidence interval contains zero. This indicates that we can not reject the null hypothesis that the different supplement types have no effect on tooth length.

```{r subgroup,echo=TRUE}
ToothGrowth.doses_0.5_1.0 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
ToothGrowth.doses_0.5_2.0 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
ToothGrowth.doses_1.0_2.0 <- subset (ToothGrowth, dose %in% c(1.0, 2.0)) 

t.test(len ~ dose, data = ToothGrowth.doses_0.5_1.0)
```

```{r doselevel, echo=TRUE}
t.test(len ~ dose, data = ToothGrowth.doses_0.5_2.0)
```

```{r unequalvar, echo=TRUE}
t.test(len ~ dose, data = ToothGrowth.doses_1.0_2.0)
```

For all three dose level pairs, the p-value is less than 0.05, and the confidence interval does not contain zero. The mean tooth length increases on raising the dose level. This indicates that we can reject the null hypothesis, and establish that increasing the dose level leads to an increase in tooth length.

4. State your conclusions and the assumptions needed for your conclusions.

## Conclusions

Supplement type has no effect on tooth growth.
Inreasing the dose level leads to increased tooth growth.
Assumptions

The experiment was done with random assignment of guinea pigs to different dose level categories and supplement type to control for confounders that might affect the outcome.
Members of the sample population, i.e. the 60 guinea pigs, are representative of the entire population of guinea pigs. This assumption allows us to generalize the results.
For the t-tests, the variances are assumed to be different for the two groups being compared. This assumption is less stronger than the case in which the variances are assumed to be equal.
